\dedicatory{Dedicated to my past, present, and future supporters\ldots}

\begin{acknowledgements}

\addchaptertocentry{\acknowledgementname}
One year ago, when I decided to enroll as a double degree student at USI, I could have never imagined how that could change my life. First of all, I would like to thank my advisor Prof. Dr. Michele Lanza. I feel extremely lucky to have been your student. You opened up my eyes, giving me the chance to see the world from a different perspective. I was fascinated by your senior experience and by your philosophical thoughts. 

This endeavor would not have been possible without the support of my two co-advisors, Dr. Csaba Nagy and Dr. Roberto Minelli. Thank you for all the time that you dedicated to me. It was essential for the success of this work.

Words cannot express my gratitude to my family and my parents Emanuele e Vera. I have been far away from home for five years, but I always felt your unconditional love and care from you. Without your support, I would never be able to come this far. 

I’m incredibly grateful to my love Biancamaria and her wonderful family. Your warmth made me feel part of the family from the very first moment. I will never forget all the things that you've done for me.   

Thanks should also go to Andrea. Together, we strengthened ourselves through our stunning university path. 
I owe special thanks to many people that do not have anything to do with my thesis but contributed to this work by making my life much better. Thank you, Daniel, Ottavio, Federico, and Carmen, for all the cool stuff we did together, which was always the cause of fun, joy, and enthusiasm.

To my Sicilian friends Francesco, Alessio, Marco, Angelo, Daniele, Davide, and many others that are too numerous to mention: in different ways and times, you all have played an essential role in my life. 

Thanks should also go to my friends Emanuele, Davide, and Samuele. I have been playing and working with you since I was 15, when we started our Minecraft community. Thanks to you, my passion took shape and became my job in life.

Lastly, I’d like to thank myself for the determination and commitment I have always put into projects I have worked on. As my grandmother said, I was a rough diamond, but now it is time to shine. 


\end{acknowledgements}